---
title: "Geography 176A"
author: "[Amy Kou](https://amykou43.github.io/amykou43/)"
subtitle: 'Lab 04: Tesselations, Point-in-Polygon'
output:
  html_document:
    theme: journal
---

```{r, warning = FALSE, message = FALSE, echo=FALSE}
library(sf)
library(USAboundaries)
library(rmapshaper)
library(tidyverse)
library(readxl)
library(gghighlight)
library(leaflet)
library(leafpop)

```

## Question 1: 

Here we will prepare five tesselated surfaces from CONUS and write a function to plot them in a descriptive way.

```{r, warning = FALSE, message = FALSE, echo=FALSE}
#1.1
CONUS = USAboundaries::us_counties()

counties = CONUS %>%
  filter(!state_name %in% c("Alaska", "Hawaii", "Puerto Rico")) %>%
 # st_union() %>%
  st_transform(5070)

#1.2
centroid = st_centroid(counties) %>%
  st_union()

#1.3
v_grid = st_voronoi(centroid) %>% #voronoi
  st_cast() %>%
  st_as_sf() %>%
  mutate(id = 1:n())

t_grid = st_triangulate(centroid) %>% #triangulated
  st_cast() %>%
  st_as_sf() %>%
  mutate(id = 1:n())

sq_grid = st_make_grid(counties, n = c(70,50)) %>% #square grid
  st_as_sf() %>%
  st_cast() %>%
  mutate(id = 1:n())

hex_grid = st_make_grid(counties, n = c(70,50), square = FALSE) %>% #hexagonal grid
  st_as_sf() %>%
  st_cast() %>%
  mutate(id = 1:n())

#1.4
t_grid = st_intersection(t_grid, st_union(counties)) #triangulated

sq_grid = st_intersection(sq_grid, st_union(counties)) #square

hex_grid = st_intersection(hex_grid, st_union(counties)) #hexagonal

v_grid = st_intersection(v_grid, st_union(counties)) #Voronoi

#1.5

CONUS_simp = ms_simplify(st_union(counties), keep = 0.05)

plot(CONUS_simp)
mapview::npts #function

CONUS_pts = mapview::npts(CONUS) #56558
CONUS_simp_pts = mapview::npts(CONUS_simp) #161

```

```{r, warning = FALSE, message = FALSE, echo= FALSE}
#1.6
plot_tess = function(data, title){
  ggplot() +
    geom_sf(data = data, fill = "white", col = "navy", size = .2) +
    theme_void() +
    labs(title = title, caption = paste("This tesselation has:", nrow(data), "tiles" )) +
    theme(plot.title = element_text(hjust = .5, color =  "navy", face = "bold"))
}

t_grid = st_intersection(t_grid, st_union(counties)) #triangulated
plot_tess(t_grid, "triangulation coverage")

sq_grid = st_intersection(sq_grid, st_union(counties))
plot_tess(sq_grid, "Square coverage")  #square

hex_grid = st_intersection(hex_grid, st_union(counties)) #hexagonal
plot_tess(hex_grid, "Hexagonal coverage")

v_grid = st_intersection(v_grid, st_union(counties)) #Voronoi
plot_tess(v_grid, "Voronoi coverage")



#1.7: The 5 plots

plot_tess(t_grid, "triangulated coverage")
plot_tess(sq_grid, "square coverage")
plot_tess(hex_grid, "Hexagonalcoverage")
plot_tess(v_grid, "Voronoi coverage")
plot_tess(data = counties, "Original")

```



## Question 2

In this question, we will write out a function to summarize our tessellated surfaces

```{r, warning = FALSE, message = FALSE, echo= FALSE}

#2.1
sum_tess = function(data, title) {
  area = st_area(data) %>% 
    units::set_units("km2") %>%
    units::drop_units()
  
  area_df = data.frame(title, length(title), mean(area), sd(area), sum(area))
  
  return(area_df)
}

#2.2 & 2.3
tess_summary = bind_rows(
  sum_tess(counties, "Counties"),
  sum_tess(v_grid, "Voroni"),
  sum_tess(t_grid, "Triangulated"),
  sum_tess(sq_grid, "Square Grid"),
  sum_tess(hex_grid, "Hexagonal"))

#2.4
knitr::kable(tess_summary,
             caption = "Tesselation Summary",
             col.names = c("Tesselation", "Number of Features", "Mean Area", "Standard Deviation of Features", "Total Area"), format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE)

```

#2.5 
We can see that the voronoi is most similar to the original. Triangulated tessellation has most features.
Hexagonal tessellation has least features.

## Question 3

We will analyze the distributions of these dams (Q3) and their purpose (Q4) through using a point-in-polygon analysis

```{r, warning = FALSE, message = FALSE, echo= FALSE}

#3.1 

NID2019_U = read_excel("~/github/geog-176A-labs/data/NID2019_U.xlsx") %>%
  filter(!is.na(LONGITUDE), !is.na(LATITUDE)) 

sf_NID2019_U = NID2019_U %>% 
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326) %>%
  st_transform(5070)

```

```{r, warning = FALSE, message = FALSE, echo= FALSE}

#3.2
point_in_polygon = function(points, polygon, id){
  st_join(polygon, points) %>%
    st_drop_geometry() %>%
    count(.data[[id]]) %>%
    setNames(c(id, "n")) %>%
    left_join(polygon, by = id) %>%
    st_as_sf()
}

```

```{r, warning = FALSE, message = FALSE, echo=FALSE}

#3.3
counties_pip = point_in_polygon(sf_NID2019_U, counties, "geoid")
v_pip = point_in_polygon(sf_NID2019_U, v_grid, "id")
t_pip = point_in_polygon(sf_NID2019_U, t_grid, "id")
sq_pip = point_in_polygon(sf_NID2019_U, sq_grid, "id")
hex_pip = point_in_polygon(sf_NID2019_U, hex_grid, "id")

```

```{r, warning = FALSE, message = FALSE, echo=FALSE}

#3.4
plot_pip = function(data, title){
  ggplot() +
    geom_sf(data = data, aes(fill = log(n)), alpha = .9, size = .2, col = NA) +
    scale_fill_viridis_c() +
    theme_void() +
    theme(plot.title = element_text(face = "bold", color = "navy", hjust = .5, size = 20)) +
    labs(title = title,caption = paste0(sum(data$n), " dams represented"))
}

```

```{r, warning = FALSE, message = FALSE, echo=FALSE}

#3.5
plot_pip(counties_pip, "Original")
plot_pip(v_pip, "Voronoi")
plot_pip(t_pip, "Triangulation")
plot_pip(sq_pip, "Square Grid")
plot_pip(hex_pip, "Hexagon Grid")

```

#3.6
Again, the voronoi is most similar to original data. I will choose voronoi because I think it has the
best coverage. The other ones do not have as much detail.

```{r, warning = FALSE, message = FALSE, echo=FALSE}
#4.1
# I chose the ones with the most dams because they seem to be the most important issues.
unique(sf_NID2019_U$PURPOSES) %>%
  length

rec_dams = sf_NID2019_U %>%
  filter(grepl("R", sf_NID2019_U$PURPOSES) == TRUE) #recreation
r_pip = point_in_polygon(rec_dams, v_grid, "id")
  
flood_dams = sf_NID2019_U %>%
  filter(grepl("C", sf_NID2019_U$PURPOSES) == TRUE) #Flood control
f_pip = point_in_polygon(flood_dams, v_grid, "id")

fire_dams = sf_NID2019_U %>%
  filter(grepl("P", sf_NID2019_U$PURPOSES) == TRUE) #Fire protection
p_pip = point_in_polygon(fire_dams, v_grid, "id")

water_dams = sf_NID2019_U %>%
  filter(grepl("S", sf_NID2019_U$PURPOSES) == TRUE) #Water Supply
w_pip = point_in_polygon(water_dams, v_grid, "id")

```

```{r, warning = FALSE, message = FALSE, echo= FALSE}

plot_pip(r_pip, "Recreation Dams") +
  gghighlight(n > (mean(n) + sd(n)))
plot_pip(f_pip, "Flood Control Dams") +
  gghighlight(n > (mean(n) + sd(n)))
plot_pip(p_pip, "Fire protection Dams") +
  gghighlight(n > (mean(n) + sd(n)))
plot_pip(w_pip, "Fish and Wildlife Dams") +
  gghighlight(n > (mean(n) + sd(n)))

```

# Extra Credit

Map the Mississippi River System and show largest/high hazard dam in each state

```{r, warning = FALSE, message = FALSE, echo=FALSE}

rivers = read_sf("~/github/geog-176A-labs/data/majorrivers_0_0")
rivers = rivers %>%
  filter(SYSTEM == "Mississippi")

# Filter to the largest/high hazard dam in each state
dams = NID2019_U %>%
  filter(HAZARD == "H", grepl("C", PURPOSES)) %>%
  group_by(STATE) %>%
  slice_max(NID_STORAGE) 

dam_labels = dams %>%
  select("DAM_NAME", "NID_STORAGE", "PURPOSES", "YEAR_COMPLETED")

radius = dams %>% 
  mutate(radius = NID_STORAGE / 1500000) %>% 
  select(radius)

avector = as.vector(radius$radius)

leaflet(data = dams) %>% 
  addProviderTiles(providers$CartoDB) %>% 
  addCircleMarkers(color = "red", fillOpacity = 1, stroke = FALSE, popup = leafpop::popupTable(dam_labels, feature.id = FALSE), radius = avector) %>% 
  addPolylines(data = rivers) 


```
