---
title: "Geography 176A"
author: "[Amy Kou](https://amykou43.github.io/amykoku43/)"
subtitle: 'Lab 02: COVID-19 Pandemic'
output:
  html_document:
    theme: journal
---

```{r, echo = FALSE, message= FALSE}

library(tidyverse)
library(knitr)
library(readxl)
library(zoo)

url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv' 

covid = read_csv(url)

PopulationEstimates = read_excel("~/Downloads/PopulationEstimates.xls", skip = 2) %>% 
  select(fips = FIPStxt, Area_Name, pop2019 = POP_ESTIMATE_2019)

state.of.interest = "California"

```

```{r, echo = FALSE, message= FALSE}
new = covid %>%
  filter(state == state.of.interest) %>%
  group_by(county) %>%
  mutate(newCases = cases - lag(cases)) %>%
  ungroup() 

 
 
(most_new_cases = new %>%
    filter(date == max(date)) %>% 
    slice_max(newCases, n = 5) %>%
    select(county, newCases)) # table #1

knitr::kable(most_new_cases, caption = 'Most New Cases California Counties',
             col.names = c("County", "New Cases"),
             format.args = list(big.mark = ",")) #table #1

```

```{r, echo = FALSE, message= FALSE}

(most_cumulative_cases = new %>%
    filter(date == max(date)) %>% 
    slice_max(cases, n = 5) %>%
    select(county, cases)) # table #2

knitr::kable(most_cumulative_cases, caption = 'Most Cumulative Cases California Counties',
             col.names = c("County", "Cum Cases"),
             format.args = list(big.mark = ",")) #table #2
```

## Attributes of Population Estimates
names of columns
dimension: 3275 x 165
number of rows: 3275
structure: tibble/data frame


```{r, echo = FALSE, eval = FALSE, message= FALSE}

colnames(PopulationEstimates) #names of columns (They're all just numbers to 165 except first column) 
dim(PopulationEstimates) #dimensions: 3275 x 165
nrow(PopulationEstimates) #number of rows: 3275
str(PopulationEstimates) #tibble
```

```{r, echo = FALSE, message= FALSE}

jd = inner_join(new, PopulationEstimates, by = "fips")
```

```{r, echo = FALSE, message= FALSE}

dat = jd %>%
  filter(state == state.of.interest) %>%
  group_by(county) %>%
  mutate(newCasesCapita = (cases - lag(cases))/pop2019,
         casePerCapita = cases / pop2019) %>%
  ungroup() 

(most_new_cases_capita = dat %>%
     filter(date == max(date)) %>% 
    slice_max(newCasesCapita, n = 5) %>%
    select(county, newCasesCapita)) 

knitr::kable(most_new_cases_capita, caption = 'Most New Cases per capita California Counties',
             col.names = c("County", "New Cases per Capita"),
             format.args = list(big.mark = ",")) #table #1

(most_cumulative_capita = dat %>%
     filter(date == max(date)) %>% 
    slice_max(casePerCapita, n = 5) %>%
    select(county, casePerCapita))

knitr::kable(most_cumulative_capita, caption = 'Most cumulative per capita California Counties',
             col.names = c("County", "Cumulative Cases per Capita"),
             format.args = list(big.mark = ",")) #table #2
  
```


```{r}
jd %>% 
  filter(date >= max(date) - 13) %>% 
  group_by(county) %>% 
  summarize(totCases = sum(newCases) / (pop2019[1]/100000)) %>% 
  filter(totCases <=  100)
  
```
```{r}
dat %>%
  summarize(totCases = sum(cases))

dat %>%
  summarize(totnewCases = sum(newCases, na.rm = TRUE))

```
## The total number of cases is 29381287 and total number of new cases is 634918 and the total number of safe counties is 12. 

```{r, warning = FALSE, message= FALSE}
state.of.interest = c("California", "New York", "Louisiana", "Florida")

to_plot = covid %>%
  filter(state %in% state.of.interest) %>%
  group_by(state, date) %>%
  summarise(dailyCases = sum(cases)) %>%
  left_join(PopulationEstimates, 
            by = c("state" = "Area_Name")) %>% 
  group_by(state) %>%
  mutate(newCases = dailyCases - lag(dailyCases),
        roll7 = rollmean(newCases, 7, fill = NA,   align="right"))  

#Plot 1  
ggplot(to_plot, aes(x = date)) +  
  geom_col(aes(y = newCases), col = NA, fill = "#F5B8B5") +  
  geom_line(aes(y = roll7), col = "darkred", size = 1) +  
  ggthemes::theme_wsj() + 
  facet_wrap(~state) + 
  labs(title = "New Cases: COVID-19") +  
  theme(plot.background = element_rect(fill = "white"),  
        panel.background = element_rect(fill = "white"),  
        plot.title = element_text(size = 14, face = 'bold')) +  
  theme(aspect.ratio = .5)


to_plot = to_plot %>% 
  mutate(nC_pc = newCases / pop2019,
         nC_pc_roll = rollmean(nC_pc, 7, fill = NA, align = "right"))

#Plot 2
ggplot(to_plot, aes(x = date)) +  
  geom_col(aes(y = nC_pc), col = NA, fill = "#F5B8B5") +  
  geom_line(aes(y = nC_pc_roll), col = "darkblue", size = 1) +  
  ggthemes::theme_wsj() + 
  facet_wrap(~state) + 
  labs(title = "New Cases per capita: COVID-19") +  
  theme(plot.background = element_rect(fill = "yellow"),  
        panel.background = element_rect(fill = "yellow"),  
        plot.title = element_text(size = 14, face = 'bold')) +  
  theme(aspect.ratio = .6)

```

## I think the scaling made some states worse and some states better. For example, for California the slope looked a lot more prevalent in New Cases but after being scaled by its huge population, it seemed less severe represented by a smaller looking slope. Louisiana had the oppoisite effect. The slope appeared bigger after being scaled with population. I think that is because its population is small. 





